Refinements for open pNets

### 2021-07-26 2021-Refinement/paper/refinementarticle.tex
related work

\usepackage{amsmath,amssymb,mathabx}
%\usepackage{amsthm}
%\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{external}
\usepackage{hyperref}
\usepackage{nccrules}
\AtBeginDocument{\renewcommand\setminus{\smallsetminus}}

@@

The notion of refinement aims of capturing the relation between  a specification and an implementation of the same component. It is usually defined as trace inclusion or simulation \cite{Milner:1980}; this ensures that all behaviours of the implementation must be also behaviours of the specification. This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems  as it requires to reason also about unspecified behaviours.

There are some works that have been interested in the  the refinement of open systems.

The definition of refinement for open systems in terms of alternating simulation \cite{Alur:1998} and \cite{deAlfaro:2021} is originating  from the game theory \cite{deAlfaro:2003}.

@@

 \(\fvars{e}\) is the set of variables in \(e \in E\) that are not bound by any binder. An expression is closed if \(\fvars{e}=\emptyset\).
The set \(\values\) denote values which is a subset of closed terms. \(\rformulas[V]\) is the set of formulas $f$ that only use variables in $V$, i.e. the formulas such that  \(\fvars{f}\subseteq V\).

@@

Similarly to FH-bisimulation we are interested here in finding relations between states of to open automata that contain variables and holes. However here we want to biuld a refinement relation that is a simulation but also guarantees that no deadlock is introduced.
Our objective is to characterize when a relation between two states of two open automata have the desired property. Such a relation has the following signature: \( S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
We first define when a relations is deadlock reducing, before characterising a first simple refinement relation only valid to compare automata with identical holes.

\begin{definition}[Deadlock reduction]\label{def:dpwd}
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if  it satisfies the following:
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{definition}

Unfortunately, in the general case, this property is not much useful in general because it conflicts with the possibility to compose automata: The composition operator can itself introduce a deadlock.
A way to solve the conflict between deadlock reduction and the properties involving composition is to only consider a composition that do not introduce deadlocks.

Before characterising a composition which does not introduce deadlocks we need to be able to tell which states are reachable.
\begin{definition}[Reachability]
For any open automata \(A = \OAg\), a reachability predicate \(\reach{A}: S \to \rformulas[V]\) is any predicate on states that is preserved accross transitions, and valdi on initial state:
\[\sigma_0 \vdash \reach{A}\mpar{s_0}\quad\land\quad\forall t = \OTg \in T, \fvars{t} \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}\]
\end{definition}
Reachability takes into account all paths, and potentially over-approximate the reachable configurations and must be representable as a formula.
% LUDO HERE
\begin{definition}[Non-locking composition, intuitive definition]
Let \(A_i = \OAg[i], 0 \leq i \leq n\).
The composition \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-locking if \(A\) has a reachability predicate satisfying:
From reachable configurations, if there was a possible transition in \(A_0\) then there is a possible transition in \(A\).
Formally for \(\OAg = A\), it gives
\begin{multline*}
	\forall s = \mpar{s'_i}_{0 \leq i \leq n} \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A}\mpar{s}} \implies \\
	\mpar{\exists t_0 \in \fOT{s'_0}, \exists \nu_0: \fvars{t_0} \to \values, \sigma \uplus \nu_0 \vdash \fguard{t_0}} \implies \\
	\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}
\end{multline*}
\end{definition}
As in the previous section, the definition can be simplified into a less intuitive version:
\begin{definition}[Non-locking composition, working definition]
\(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is a non-locking composition if:
\[ \forall s = \mpar{s_0, s'} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \]
\end{definition}
\begin{lem} The intuitive and working definition of non-locking composition are equivalent. \end{lem}
We want to prove
\[ \forall s = \mpar{s_p, s_c} \in S, V \uplus \biguplus_\subbox{t_p \in \fOT{s_p}} \fvars{t_p} \vdash \reach{A}\mpar{s} \wedge \bigvee_\subbox{t_p \in \fOT{s_p}} \fguard{t_p} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \tag{WD}\label{eq:nlcWD} \]
\[ \bigsymb{\iff} \]
\begin{multline*}
	\forall s = \mpar{s_p, s_c} \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A}\mpar{s}} \implies \\
	\mpar{\exists t_p \in \fOT{s_p}, \exists \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \vdash \fguard{t_p}} \implies \\
	\mpar{\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \tag{ID}\label{eq:nlcID}
\end{multline*}

>>>>

\usepackage{amsmath,amssymb}
%\usepackage{amsthm}
%\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{external}
\usepackage{hyperref}
\usepackage{nccrules}
\AtBeginDocument{\renewcommand\setminus{\smallsetminus}}

@@

The notion of refinement aims of capturing the relation between  a specification and an implementation of the same component. It is usually defined as trace inclusion or simulation \cite{Milner:1980}; this ensures that all behaviours of the implementation must be also behaviours of the specification. This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.

There are some works that have focused on  the refinement of open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
\cite{GayH:2005,BravettiZ:2021}. Such refinement is  instead based  on interface-oriented approach, it allows the expression more internal choices and less external choices. The refinement of open automata is defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}, which is based on a process-oriented approach. Alternating simulation that is originating from the game theory \cite{deAlfaro:2003} allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement
of automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions).
The result of the composition of automata may
contain illegal states, where one automaton issues an output that is not acceptable as input in the other one. An optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, they can be composed together. Although our approach to design refinement differs from that of the above mentioned \cite{deAlfaro:2021},  because it is based on the notion of simulation, it has some commonalities:  our formalism is based on transition systems and on  optimistic approach to composition.

@@

 \(\fvars{e}\) is the set of variables in \(e \in E\) that are not bound by any binder. An expression is closed if \(\fvars{e}=\emptyset\).
The set \(\values\) denote values which is a subset of closed terms.

@@

deadlock and reachability
